---
title: "Assignment 1"
author: "David Nguyen 300584723"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

### a.

Numerical - Continuous - numeric - double

### b. 

Categorical - Logical - logical - logical

### c.

Numerical - discrete - numeric - integer

### d.

Categorical - Nominal - character - character

### e.

Categorical - Ordinal - ordered factor - integer

## Q2.

```{r}
y <- c("Earth"=0, "Water"=-3, "Fire"=2)
y
```
### a.

```{r}
y[y==2]
```

### b.

```{r}
names(y[y==2])
```

### c.
```{r}
a <- c(3,4,1)
a
z <- matrix(c(0,3,0,9,2,1), nrow = 2, ncol = 3)
z

```
### d.
```{r}
max(a)
```

## Q3.

### a.

```{r, error=TRUE}
standardize_column <- function(mat, col_index){
    mean <- mean(mat[,col_index], na.rm = TRUE)
    std <- sd(mat[,col_index], na.rm = TRUE)
    z <- ((mat[,col_index]-mean)/std)
    print(z)
}

```
### b.
```{r, error=TRUE}
mat <- matrix(c(12,1,4,3,NA,9,NA,5,5), nrow=3)
test <- standardize_column(mat,2) #Testing on column 2
```

### c.
```{r, error=TRUE}
standardize_column <- function(mat, col_index){
  # Modified function ------------------------------
  if (col_index>=ncol(mat) ||  #check if index greater than n.o cols
      col_index <= 0||  # check if col_index negative or zero  
      col_index != as.integer(col_index)){ #check if col_index is decimal index condition
    stop("col must be a valid index of mat")
  }
  else {
    mean <- mean(mat[,col_index], na.rm = TRUE)
    std <- sd(mat, na.rm = TRUE)
    z <- ((mat[,col_index]-mean)/std)
    print(z)
  }
}
```

### d.
```{r, error=TRUE}
standardize_column(mat,.3)
standardize_column(mat, -1)

```
## Q4.

### a.
```{r}
stars <- read.csv("star_dataset_deepu_kaggle.csv")
```
### b.
```{r}
head(stars,2)
```
### c.
```{r}
print(paste("Number of rows of stars:",nrow(stars)))
print(paste("Number of columns of stars:",ncol(stars)))
```
### d.
```{r}
subset_stars <- stars[stars$Star.color == "Blue" |stars$Star.color == "White"|stars$Star.color == "Orange",]
subset_stars
```
### e.
```{r}
print(class(subset_stars$Star.color))
```
We can see that Star.color is being treated as character instead of factor.

### f.
``` {r}
subset_stars$star_colours <- factor(subset_stars$Star.color, levels = c("Blue", "White", "Orange"))
```

### g.
```{r}
#Create a frequency table
color_counts <- table(subset_stars$star_colours)
color_counts
# Create a bar plot
barplot(color_counts, main = "Frequency of Star Colors", 
        xlab = "Star Color", ylab = "Number of Stars", col = c("navy", "grey", "orange2"),
        legend.text = levels(subset_stars$star_colours))
```

## Q5.
x=y assign y to x

x==y check if x equals to y, it can return either TRUE or FALSE

x<-y same with the case of x=y, it will asign the value of y to x
but the "=" sign is more commonly used in function calls rather than
assigning to prevent from facing complex expressions

x<(-y) checks if x is smaller than the negative value of y

x^(-y) shows the value of x with the power of negative value of y

## Q6.
### a.
```{r}
library("palmerpenguins")
penguins
hist(penguins$body_mass_g, main="Distribution of Penguins Body Mass", xlab = "Penguins Weight (g)", col = "lightgreen")
```

### b.
```{r}
plot(penguins$flipper_length_mm, penguins$bill_length_mm,
     main = "Scatter Plot of Flipper Length and Bill Length", col = "lightblue", xlab = "Fliiper length (mm)",
     ylab ="Bill length (mm)",pch=16)

```

### c.

```{r}

```

```{r}
library("palmerpenguins")
barplot(table(penguins$species, penguins$island), beside=TRUE,
        xlab = "Island", ylab="Frequency", legend.text = TRUE, col = c("pink","khaki","lightgreen"))
```


### d.

```{r}
boxplot(penguins$body_mass_g~penguins$sex, main="Body Mass by Sex",
        col = "blue",xlab = "Gender",ylab = "Weight (g)")
```

## Q7.
### Line 2

```{r, eval = FALSE}
# Define x
x <= c(3,0,5,6,NA,3,1)
#### FINDING: it is supposed to be 'x <-' not 'x <='
```

```{r}
### CORRECTED CODE:
x <- c(3,0,5,6,NA,3,1)
```

### Line 4 
```{r, eval = FALSE}
# Store the first five entries of x as a new object called y
y <- x[0:4]
### FINDING: R indices are 1-based which means the vector indices start from 1 not 0 like other languages. Therefore, in order to
# store the first five entries, we need to rewrite the code from '[0:4] to [1:5]'
```


```{r}
### CORRECTED CODE:
y <- x[1:5]
```
### Line 6
```{r, eval =FALSE}
# Calculate the maximum value of y after excluding NA entries
max_y <- max(y, na.remove=TRUE)
### FINDING: the second argument is supposed to be 'na.rm' not 'na.remove'
```

```{r}
### CORRECTED CODE:
max_y <- max(y,na.rm = TRUE)
```

### Line 8
```{r}
# Fetch the entries of y that are bigger than 2
y[y >= 2]
### FINDING: The sign >= means 'bigger or equal to 2', we need to use '>' instead 
```

```{r}
### CORRECTED CODE:
y[y>2]
```